---
title: "Google Summer of Code Contributions"
author: "Lorenz Walthert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Google Summer of Code Contributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vigette aims to summarize the contributions made to styler through 
Google Summer of Code 2017.
```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
knitr::knit_engines$set(list(
  styler = function(options) {
    options$comment <- ""
    knitr::engine_output(
      options,
      
      c("# Before", options$code),
      c("# After", styler::style_text(options$code))
    )
  }
))

```

As a warm-up, we want to give a few examples of how styler styles code. It's
possible to use different levels of invasiveness, as described in the help
file for the style guide we implemented, which is the [tidyverse style guide](http://style.tidyverse.org/index.html). For example, we can limit 
ourselfs to styling just spacing information.
```{r}
library("styler")
library("magrittr")
style_text("a=3; 2", scope = "spaces") %>% cat()
```

Or, on the other extreme of the scale, styling spaces, indention, line breaks and tokens:
```{r}
style_text("a=3; 2", scope = "tokens") %>%
  cat(sep = "\n")
```
Another option that is helpful to determine the level of invasiveness is
`strict`. If set to `TRUE`, spaces and line breaks after or before tokens are
set to either zero or one. However, in some situations that might be undesirable,
as the following example shows
```{r}
style_text(
  "data_frame(
     small  = 2 ,
     medium = 4,#comment without space
     large  = 6
   )", strict = FALSE) %>%
  cat(sep = "\n")
```

I hope these examples showed you some of the options available in styler. Let
us for now focus on a configuration with `strict = TRUE` and `scope = "tokens"` 
and therefore only show code before and after styling. 

styler can distinguish unary operators and other math tokens

```{styler}
1++1-1-1/2
```

It can also format complicated expressions that involve line breaking and 
indention based on both brace expressions and operators
```{styler}
if (x >3) {stop("this is an error")} else {
c(there_are_fairly_long,
1 / 33 * 
2 * long_long_variable_names)%>% k(

) }
```

Lines are broken after `(` if a function calls spans over multiple lines
```{styler}
do_a_long_and_complicated_fun_cal("which", has, way, to, 
                              "and longer then lorem ipsum in its full length"
                              )
```

Styler converges `=` assignment, Single quotes if necessary, adds braces to
function calls in pipies
```{styler}
one= 'one string'
two= "one string in a 'string'"
a %>%
  b %>%
  c

```

Function declarations are indented if multi-line
```{styler}
my_fun <- function(x, 
y, 
z) {
  just(z)
}
```

Styler can also deal with tidyeval syntax
```{styler}
mtcars %>%
  group_by(!!!my_vars)
```


# High-level overview
The author 

* opened over [60](https://github.com/krlmlr/styler/issues?utf8=✓&q=is%3Aopen%20is%3Aissue%20author%3Alorenzwalthert%20created%3A%3C2017-08-31%20%20created%3A%3E2017-06-01%20) issues and closed the majority of them with his own pull requests.
* created more than [60](https://github.com/krlmlr/styler/pulls?q=is%3Apr+author%3Alorenzwalthert+created%3A%3C2017-08-31++created%3A%3E2017-06-01) pull requests to
  master of which the vast majority was merged.

The activity during GSOC can be grouped into the following steps:

* Replicating the outcomes that were achieved with flat styling with nested 
  styling. Required implementeation of the visitor as well as different 
  serialization method.
* Adding functionality that was only possible with the nested styling such 
  as indention based on operators (`+`, `%>%` etc.), braces `{` and `(`. 
  the use of raw indention, allowing for token-dependent re-indention.
* Infrastructure changes such as serialized testing, improve performance, 
  create mutli-line attribute, code refactoring. Partially
  flattening out the nested parse table.
* Extending the functionality that was not there with the flat approach. 
  Examples include styling multiple expressions, token replacement (e.g. `<-` for 
  `=`), line break rules (invasive and non-invaive), token insertion (`()` in 
  function calls within pipes), RStudio Addins, styling of comments, styling of 
  a single file, public api redesign, spacing around more operators (colons, 
  commas, infix operators, bang-bang).

The list above is not comprehensive and covers only the large contributions.
Always along with coding, documentation and tests were provided. With regard
to the initial [proposal](https://docs.google.com/document/d/1V7ZWOVjEqM6-vc7WVxePUOHHc48HJwBKh8gAwIFALJ4/edit), we can say that steps one to four were completed
to a large extend and a few points under step 5 have been implemented too.
The tidyverse style guide does not cover all formatting cases, so 
in particular for function calls, there are various cases uncovered and 
a corresponding [issue](https://github.com/krlmlr/styler/issues/125) was opened
in the styler repo.
Also, it was not possible in this project to implement the rule that forces
lines to be less than 80 characters long. We leave this to [lintr](https://github.com/jimhester/lintr). Known remaining problems are 
well-documented on the [issue page]([issue](https://github.com/krlmlr/styler/issues) of the main repository on GitHub.

It is nice to see that `styler` already got some traction from the community.
The maintainer of the package [reprex](https://github.com/tidyverse/reprex) 
plans to [add styler as a dependency](https://github.com/tidyverse/reprex/issues/94).


